---
title: "Reproducible Research - Peer Assessment 1"
output: html_document
---
###Loading and preprocessing the data

```{r}
 Step_Activity <- read.csv("activity.csv")
```

### Q1: What is mean total number of steps taken per day?

```{r}
#Sum steps for each date
StepsperDay<-aggregate(steps~date, Step_Activity,sum)
#Plot histogram
hist( StepsperDay$steps,breaks=20,col="blue",xlab ="No of Steps",main="Steps ~ Date")
# Finding MEAN and MEDIAN
mean(StepsperDay$steps)
median(StepsperDay$steps)
```

### Q2: What is the average daily activity pattern?
####   a)Time series plot:  5-minute interval ~ the average steps taken, across all days (y-axis)
```{r}
StepsPerInterval<-aggregate(steps~interval,Step_Activity,sum)
plot(StepsPerInterval$interval,StepsPerInterval$steps,  ylab="Steps", xlab="5 Min Intervals", pch=".")
lines(StepsPerInterval$interval,StepsPerInterval$steps)
```

####   b)Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
MaxStepInterval <- StepsPerInterval[StepsPerInterval$steps==max(StepsPerInterval$steps),1]
MaxStepInterval
```

### Q3: Inputing missing values
####   a)Total number of missing values in the dataset
```{r}
NAVector<-is.na( Step_Activity)        
NARows <-  Step_Activity[NAVector, ]   # Creating new data set with NA rows alone
length( NARows$steps)
```
####    b)Create a new dataset that is equal to the original dataset but with the missing data filled in
#####STRATEGY: Take out all rows for each 5 min interval and replace NA with steps mean of that interval
```{r}
NonNAData<- Step_Activity
Values<-unique(NonNAData$interval)         # finding no of unique intervals
#copy each interval data into a temporary dataset and replace NA with mean of that interval
for(k in Values){                                    
TempData<-NonNAData[NonNAData$interval==k,]
NAVector<-is.na(TempData)
TempData[NAVector[,1],1] <- mean(TempData[!NAVector[,1],1])
NonNAData[NonNAData$interval==k,] <- TempData
}
```

####   c)Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day
```{r}
#Find sum of steps per Day for original and new data
StepsPerDayOld<-aggregate(steps~date,Step_Activity,sum)
StepsPerDayNew<-aggregate(steps~date,NonNAData,sum)
#PLOT histogram with new data obtained by replacing NAs with means
hist(StepsPerDayNew$steps,breaks=61,col="blue",xlab ="No of Steps",main="Steps ~ Date")
#OLD MEAN and MEDIAN
mean(StepsPerDayOld$steps)
median(StepsPerDayOld$steps)
#NEW MEAN and MEDIAN
mean(StepsPerDayNew$steps)
median(StepsPerDayNew$steps)
```

### Q4: Are there differences in activity patterns between weekdays and weekends?

#####  a)Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a                 given date is a weekday or weekend day
```{r}
 Wday_Wend <- (weekdays(as.Date(Step_Activity$date)) %in% c('Saturday','Sunday'))
 Day <- 1: length(Step_Activity$date)
 TempData<-Step_Activity
 TempData<-cbind(TempData,Day)
 TempData[!Wday_Wend,4]<-"WEEKDAY"
 TempData[Wday_Wend,4]<-"WEEKEND"

```

####   b)Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)
```{r}
 Totals<-aggregate(steps~interval+Day,TempData,sum)
 library(lattice)
 xyplot(steps~interval|Day,data=Totals,type="a"
  ,layout=c(1,2),xlab="5 Min Intervals"
  , ylab= "No Of Steps")
```